---
title: "GCAT Circular Codes Tutorial"
author: Martin Starman, Markus Gumbel
date: "`r format(Sys.time(), '%d. %B %Y')`"
params:
  devel: TRUE
output:
 html_document:
  toc: true
  number_sections: true 
---

_under construction_

```{r include=params$devel, eval=params$devel, echo = F,  message=FALSE}
devtools::load_all(".")
```

```{r include=!params$devel, eval=!params$devel}
library(gcatcirc)
```

# Let's go

Let us start with general code Functions. With the function `is_code(tuples)` you can check whether  a set of tuples is a code. Suppose X is a set of words. This set X is a code if and only if all possible sequences have at most one decomposition into words of X.

```{r}
X = gcatbase::code(c("ACG", "CGG", "AC"))
is_code(X)
```
If X is not a code:
```{r}
X = gcatbase::code(c("ACG", "CGG", "AC", "GCGG"))
is_code(X)
```
Then the function `all_ambiguous_sequences(tuples)` returns all ambiguous sequences:
```{r}
print(all_ambiguous_sequences(X))
```
With the function `circular_shift(tuples, sh)` you can circular permutate a code, i.e. let X={123, 332}, then c.shift(2) results in {312, 233}

```{r}
code <- gcatbase::code(c("ACG", "CGG", "AC"))
print(circular_shift(code, 2))
```

# Circular code tools

The function `is_code_circular(tuples)` checks if a code is circular.

```{r}
code <- gcatbase::code(c("ACG", "CGG", "AC"))
is_code_circular(code)
```
Other properties in the context of circularity can be tested with the following functions. The description of the properties can be looked up in the docs.

```{r}
code <- gcatbase::code(c("ACG", "CGG", "AC"))
cat("k-graph: ",  get_k_graph_circular(code), "\n")
cat("k-circular: ", get_exact_k_circular(code), "\n")
cat("is: k-circular: ", is_code_cn_circular(code), "\n")
cat("is: comma free: ", is_code_comma_free(code), "\n")
cat("is: strong comma free: ", is_code_strong_comma_free(code), "\n")
```

# C3 codes

`c3_codes()` Returns a list of maximal self-complementary C3 codes. Without paramet it returns all C3 codes:

```{r}
n = c3_codes()
print(length(n))
```

There is also a function which allows direct access to each $C_3$-gcatbase::gcat.code:

```{r}
X0 = c3_code(23)
print(X0)
summary(X0)
```

```{r}
print(gcatbase::amino_acids(X0))
```

Equivalence classes

Let us determine the equivalence class for $X_0$:
```{r}
ecl = c3_equiv_class(23)
print(ecl)
```

Now we want to see all other $C_3$-codes of this class:
```{r}
l = c3_in_class(ecl)
print(l)
```

Here are the codons:

```{r}
cl = c3_codes(l)
print(cl)
```

How many amino acids do these codes cover?

```{r}
al = sapply(l, function(i) length(unique(amino_acids(c3_code(i)))))
print(al)
```

# Graph tools

You can plot a responding graph with the following chunk
```{r}
code <- gcatbase::code(c("ACG", "CGG", "AC"))
G <- get_representing_graph(code, TRUE, TRUE)
plot(G)
```

To plot a component simply run;

```{r}
code <- gcatbase::code(c("ACGG", "CGGC", "AC"))
G <- get_component_of_representing_graph(code, 2, TRUE, TRUE)
plot(G)
```

The function `plot_representing_graph()` and `plot_component_of_representing_graph()` are simple shortcuts to use `igraph::tkplot()`

To get the longest path within a representing graph you can run:
```{r}
code <- gcatbase::code(c("ACGG", "CGGC", "AC"))
get_longest_paths(code)
```
The same works for all cyclic paths with:

```{r}
code <- gcatbase::code(c("ACGG", "GGGC", "GCAC", "GGAC", "AC"))
get_cyclic_paths(code)
```