---
title: "GCAT Circular Codes Tutorial"
author: Martin Starman, Markus Gumbel
date: "`r format(Sys.time(), '%d. %B %Y')`"
params:
  devel: TRUE
output:
 html_document:
  toc: true
  number_sections: true 
---

_under construction_

```{r include=params$devel, eval=params$devel}
devtools::load_all(".")
```

```{r include=!params$devel, eval=!params$devel}
library(gcatcirc)
```

```{r}
library(gcatbase)
```

# Let's go

```{r}
print(gcatcirc::hellogcatcirc())
```

# C3 codes

`all_c3_codes` gives all $C_3$ codes:

```{r}
n = length(all_c3_codes())
print(n)
```

There is also a function which allows a direct access to each $C_3$-code:

```{r}
X0 = c3_code(23)
print(X0)
summary(X0)
```

```{r}
print(amino_acids(X0))
```

Equivalence classes

TODO: Maybe we better pass as argument the code itself not 23.

Let us determine the equivalence class for $X_0$:
```{r}
ecl = c3_equiv_class(23)
print(ecl)
```

Now we want to see all other $C_3$-codes of this class:
```{r}
l = c3_in_class(ecl)
print(l)
```

Here are the codons:

```{r}
cl = lapply(l, function(i) c3_code(i))
print(cl)
```

How many different amino acids do these code cover?

```{r}
al = sapply(l, function(i) length(unique(amino_acids(c3_code(i)))))
print(al)
```
